---
title: "hw9_531"
author: "mengjiao"
date: "April 3, 2016"
output: html_document
---

###1. Assessing and improving algorithmic paramters
(1)cooling.fraction.50
If we cooled quickly, the convergence plots may converge faster or we may miss the maximum log likelihood. If we cooled slowly, it may do not coverge in a given iteration.

(2)mif_iterations
300 is large enough for the parameters and loglik to converge. It only needs about 150 to 200 iterations to converge from the diagnosis of the convergence.

(3)Different mif runs 
Differenct initial parameters lead to very large inconsistency in the convergent parameter. Different beta initilization lead to similar choice. But different mu_I and rho lead to very different parameters. However, loglik seems to converge. Therefore, it may show that it is a flat likelihood, since different parameters leads to similar likelihood.

(4)Particle filter
From the last iteration, it shows at time 12, the effective sample size is rather small. May we need to increase the number of particles.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```


```{r}
summary(cars)
```



```{r, echo=FALSE}
plot(cars)
```


