---
title: "STATS531 HW10"
author: "Xiaoyue Liu    UMID:28589009    xiaoyliu@umich.edu"
date: "April 11, 2016"
output:
  html_document:
    fig_caption: true
    theme: flatly
    toc: yes
    toc_depth: 2
    number_sections: true
    pandoc_args: [
      "--number-offset=10"
    ]
---

\newcommand\prob{\mathbb{P}}
\newcommand\E{\mathbb{E}}
\newcommand\var{\mathrm{Var}}
\newcommand\cov{\mathrm{Cov}}
\newcommand\loglik{\ell}
\newcommand\R{\mathbb{R}}
\newcommand\data[1]{#1^*}
\newcommand\params{\, ; \,}
\newcommand\transpose{\scriptsize{T}}
\newcommand\eqspace{\quad\quad}
\newcommand\myeq[1]{\eqspace \displaystyle #1}
\newcommand\lik{\mathscr{L}}
\newcommand\loglik{\ell}
\newcommand\profileloglik[1]{\ell^\mathrm{profile}_#1}
\newcommand\ar{\phi}
\newcommand\ma{\psi}
\newcommand\AR{\Phi}
\newcommand\MA{\Psi}
\newcommand\ev{u}
\newcommand\given{{\, | \,}}
\newcommand\equals{{=\,}}
\newcommand\matA{\mathbb{A}}
\newcommand\matB{\mathbb{B}}
\newcommand\matH{\mathbb{H}}
\newcommand\covmatX{\mathbb{U}}
\newcommand\covmatY{\mathbb{V}}


-------

--------
```{r knitr-opts,include=FALSE,purl=FALSE,cache=FALSE}
prefix <- ''
library(knitr)
opts_chunk$set(
  progress=TRUE,
  prompt=FALSE,tidy=FALSE,highlight=TRUE,
  strip.white=TRUE,
  warning=FALSE,
  message=FALSE,
  error=FALSE,
  echo=TRUE,
  cache=TRUE,
  cache.extra=rand_seed,
  results='markup',
  fig.show='asis',
  size='small',
  fig.lp="fig:",
  fig.path=paste0("figure/",prefix,"-"),
  cache.path=paste0("cache/",prefix,"-"),
  fig.pos="h!",
  fig.align='center',
  fig.height=4,fig.width=6.83,
  dpi=300,
  dev='png',
  dev.args=list(bg='transparent')
  )

```
```{r opts,include=FALSE,cache=FALSE}
options(
  keep.source=TRUE,
  stringsAsFactors=FALSE,
  encoding="UTF-8"
  )
```

```{r prelims,echo=F,cache=F}
set.seed(594709947L)
require(ggplot2)
theme_set(theme_bw())
require(plyr)
require(reshape2)
require(foreach)
require(doParallel)
require(pomp)
stopifnot(packageVersion("pomp")>="0.69-1")
```

* My data is from Project Tycho.


```{r read_data}
CA <- read.csv("HEPATITIS_A_Cases_CALIFORNIA.csv",skip=2,header=T)
CA <- CA[CA$YEAR>=1966,]
CA$Time <- CA$YEAR+CA$WEEK/52
plot(CA$Time,CA$CALIFORNIA,main="Hepatitis A cases in California from 1966 to 2011",type='l')

MI <-read.csv("HEPATITIS_A_Cases_MICHIGAN.csv",skip=2,header=T)
MI <- MI[MI$YEAR>=1966,]
MI$Time <- MI$YEAR+MI$WEEK/52
plot(MI$Time,MI$MICHIGAN,main="Hepatitis A cases in Michigan from 1966 to 2011",type='l')
blowfly <- read.csv("blowfly4.csv",skip=3, header=T)
```

blowfly <- read.csv("blowfly4.csv",skip=3, header=T)
